Dashboards for stock assessments: Getting started

### 2019-10-18 analysis/dashboards/getting_started.Rmd
Draft a plotly basics page to the getting_started dashboard

@@

The `flexdashboard` package can be used to render groups of related text, figures and tables into a dashboard. Using a combination of `markdown` syntax and `R` code (i.e. R markdown), this package facilitates a wide range of layout options and each component can include output from packages such as `plotly`, `leaflet`, `ggplot2`, and so on. The package also integrates nicely with `shiny` and `crosstalk`, providing options for increasing the interactivity of the dashboard. Some layout and component options are highlited below.

>>>>

The `flexdashboard` package can be used to render groups of related text, figures and tables into a dashboard. Using a combination of `markdown` syntax and `R` code (i.e. R markdown), this package facilitates a wide range of layout options and each component can include output from packages such as `plotly`, `leaflet`, `ggplot2`, and so on. The package also integrates nicely with `shiny` and `crosstalk`, providing options for increasing the interactivity of the dashboard. Some layout and component options are highlighted below.

plotly {data-navmenu="Basics" data-orientation=columns}
=======================================================================

Column {data-width=250}
-------------------------------------

There are two ways to produce interactive graphics using `plotly`:

1. Using the function `ggplotly` to convert plots from `ggplot` to `plotly` objects
2. Using the native syntax of the `plotly` package

Both options follow the layered grammar of graphics, which is a generic tool for concisely describing the components of a graphic, such as the:

- Data and mappings (aesthetics)
- Geometric objects (points, lines, polygons)
- Scales (color, size, shape, axes)

Column {data-width=400 .tabset}
-------------------------------------

### ggplotly

Using the `ggplotly` function, a user can construct a plot using `ggplot2` and supply the plot object to `ggplotly` to convert it to an interactive `plotly` graphic:

```{r, echo = TRUE, eval = FALSE}
p <- ggplot(data = iris, 
            mapping = aes(x = Sepal.Length, 
                          y = Petal.Length, 
                          color = Species)) +
  geom_point() +
  scale_color_manual(values = viridis::viridis(3))
ggplotly(p)
```

#### Data and mapping

Here, the `ggplot` function is used to set-up the base layer of the plot as well as the aesthetics. That is, `iris` data are supplied to the function and specific columns in the `iris` data-set are mapped to `x`, `y` and `color` aesthetics.

#### Geometric objects

Next points are added using the `geom_point` function. Many other geometries can be used to visually represent the data, such as lines (`geom_line`), bars (`geom_bar`), text (`geom_text`).

#### Scales

Finally, colors were modified using `scale_color_manual`. While the colors are manually defined here, `ggplot2` includes built in color scales such as `scale_color_grey`. Scales used for the `x` and `y` aesthetics can also be modified using functions such as `scale_x_log10`.

### Output

```{r}
p <- ggplot(data = iris, 
            mapping = aes(x = Sepal.Length, 
                          y = Petal.Length, 
                          color = Species)) +
  geom_point() +
  scale_color_manual(values = viridis::viridis(3))
ggplotly(p)
```

Column {data-width=400 .tabset}
-------------------------------------

### plot_ly

Using functions from the `plotly` package, `plotly` plots can be directly created using syntax similar to `ggplot2`:

```{r, echo = TRUE, eval = FALSE}
plot_ly(data = iris, 
        x = ~Sepal.Length, 
        y = ~Petal.Length, 
        color = ~Species,
        colors = viridis::viridis(3)) %>% 
  add_markers()
```

#### Data and mapping

The `plot_ly` function is analogus to the `ggplot` function from `ggplot2` and, like `ggplot`, this function sets-up the base layer of the plot. Again, the `iris` data is supplied and specific columns are mapped to `x`, `y` and `color` aesthetics.

#### Geometric objects

Under `plotly`, geometries are added using functions with the `add_` prefix rather that the `geom_` prefix of used in `ggplot2`. Points are added here using `add_markers` and, like `ggplot2`, several other geometries can be specified, such as lines (`add_lines`), bars (`add_bars`), text (`add_text`).

#### Scales

Unlike `ggplot2`, scales are not modified using specific functions with a `scale_` prefix, rather, scales are specified in the `plot_ly` function call itself or via the `layout` function. Here, colors are specified using the `colors` argument in the `plot_ly` function.

### Output

```{r}
plot_ly(data = iris, 
        x = ~Sepal.Length, 
        y = ~Petal.Length, 
        color = ~Species,
        colors = viridis::viridis(3)) %>% 
  add_markers()
```

@@
